---
title: "Machine Learning-Heart Risk"
author: "Glen Lin"
date: '2024-5-29'
output: html_document
---

### Define function

#### Confusion Matrix

```{r message=FALSE, warning=FALSE}
 draw_confusion_matrix <- function(cm) {

  layout(matrix(c(1,1,2)))
  par(mar=c(2,2,2,2))
  plot(c(100, 345), c(300, 450), type = "n", xlab="", ylab="", xaxt='n', yaxt='n')
  title('CONFUSION MATRIX', cex.main=2)

  # create the matrix 
  rect(150, 430, 240, 370, col='#678197')
  text(195, 435, "Heart Attack Risk", cex=1.2)
  rect(250, 430, 340, 370, col='#a2d2df')
  text(295, 435, "Less Risk", cex=1.2)
  text(125, 370, 'Predicted', cex=1.3, srt=90, font=2)
  text(245, 450, 'Actual', cex=1.3, font=2)
  rect(150, 305, 240, 365, col='#678197')
  rect(250, 305, 340, 365, col='#a2d2df')
  text(140, 400, "Heart Attack Risk", cex=1.2, srt=90)
  text(140, 335, "Less Risk", cex=1.2, srt=90)

  # add in the cm results 
  res <- as.numeric(cm$table)
  text(195, 400, res[4], cex=1.6, font=2, col='white')
  text(195, 335, res[3], cex=1.6, font=2, col='white')
  text(295, 400, res[2], cex=1.6, font=2, col='white')
  text(295, 335, res[1], cex=1.6, font=2, col='white')

  # add in the specifics 
  plot(c(100, 0), c(100, 0), type = "n", xlab="", ylab="", main = "DETAILS", xaxt='n', yaxt='n')
  text(10, 85, names(cm$byClass[1]), cex=1.2, font=2)
  text(10, 70, round(as.numeric(cm$byClass[1]), 3), cex=1.2)
  text(30, 85, names(cm$byClass[2]), cex=1.2, font=2)
  text(30, 70, round(as.numeric(cm$byClass[2]), 3), cex=1.2)
  text(50, 85, names(cm$byClass[5]), cex=1.2, font=2)
  text(50, 70, round(as.numeric(cm$byClass[5]), 3), cex=1.2)
  text(70, 85, names(cm$byClass[6]), cex=1.2, font=2)
  text(70, 70, round(as.numeric(cm$byClass[6]), 3), cex=1.2)
  text(90, 85, names(cm$byClass[7]), cex=1.2, font=2)
  text(90, 70, round(as.numeric(cm$byClass[7]), 3), cex=1.2)

  # add in the accuracy information 
  text(30, 35, names(cm$overall[1]), cex=1.5, font=2)
  text(30, 20, round(as.numeric(cm$overall[1]), 3), cex=1.4)
}  

```

#### ROC Curve 

```{r message=FALSE, warning=FALSE}
specificity=function (actuals, predictedScores, threshold = 0.001) 
{
    predicted_dir <- ifelse(predictedScores < threshold, 0, 1)
    actual_dir <- actuals
    no_without_and_predicted_to_not_have_event <- sum(actual_dir != 
        1 & predicted_dir != 1, na.rm = T)
    no_without_event <- sum(actual_dir != 1, na.rm = T)
    return(no_without_and_predicted_to_not_have_event/no_without_event)
}
sensitivity=function (actuals, predictedScores, threshold = 0.001) 
{
    predicted_dir <- ifelse(predictedScores < threshold, 0, 1)
    actual_dir <- actuals
    no_with_and_predicted_to_have_event <- sum(actual_dir == 
        1 & predicted_dir == 1, na.rm = T)
    no_with_event <- sum(actual_dir == 1, na.rm = T)
    return(no_with_and_predicted_to_have_event/no_with_event)
}
AUROC=function (actuals, predictedScores) 
{
    numrow = length(seq(max(predictedScores, 1, na.rm = T), (min(predictedScores, 
        0, na.rm = T) - 0.02), by = -0.02))
    df <- as.data.frame(matrix(numeric(numrow * 2), ncol = 2))
    names(df) <- c("One_minus_specificity", "sensitivity")
    rowcount = 1
    getFprTpr <- function(actuals, predictedScores, threshold = 0.001 ){
        return(list(1 - specificity(actuals = actuals, predictedScores = predictedScores, 
            threshold = threshold), sensitivity(actuals = actuals, 
            predictedScores = predictedScores, threshold = threshold)))
    }
    for (threshold in seq(max(predictedScores, 1, na.rm = T), 
        (min(predictedScores, 0, na.rm = T) - 0.02), by = -0.02)) {
        df[rowcount, ] <- getFprTpr(actuals = actuals, predictedScores = predictedScores, 
            threshold = threshold)
        rowcount <- rowcount + 1
    }
    df <- data.frame(df, Threshold = seq(max(predictedScores, 
        1, na.rm = T), (min(predictedScores, 0, na.rm = T) - 
        0.02), by = -0.02))
    auROC <- 0
    for (point in c(2:nrow(df))) {
        x1 <- df[point - 1, 1]
        x2 <- df[point, 1]
        y1 <- df[point - 1, 2]
        y2 <- df[point, 2]
        rect_x <- x2 - x1
        rect_y <- y1
        rect_area <- rect_x * rect_y
        triangle_area <- rect_x * (y2 - y1) * 0.5
        currArea <- rect_area + triangle_area
        auROC <- auROC + currArea
    }
    totalArea <- (max(df[, 1]) * max(df[, 2]))
    return(auROC/totalArea)
}

ROC.plot.rv<-function (actuals, predictedScores) 
{
    One_minus_specificity <- Threshold.show <- NULL
    numrow = length(seq(max(predictedScores, 1, na.rm = T), (min(predictedScores, 
        0, na.rm = T) - 0.02), by = -0.02))
    df <- as.data.frame(matrix(numeric(numrow * 2), ncol = 2))
    names(df) <- c("One_minus_specificity", "sensitivity")
    rowcount = 1
    getFprTpr <- function(actuals, predictedScores, threshold = 0.5) {
        return(list(1 - specificity(actuals = actuals, predictedScores = predictedScores, 
            threshold = threshold), sensitivity(actuals = actuals, 
            predictedScores = predictedScores, threshold = threshold)))
    }
    for (threshold in seq(max(predictedScores, 1, na.rm = T), 
        (min(predictedScores, 0, na.rm = T) - 0.02), by = -0.02)) {
        df[rowcount, ] <- getFprTpr(actuals = actuals, predictedScores = predictedScores, 
            threshold = threshold)
        rowcount <- rowcount + 1
    }
    AREAROC <- AUROC(actuals = actuals, predictedScores = predictedScores)
    df <- data.frame(df, Threshold = seq(max(predictedScores, 
        1, na.rm = T), (min(predictedScores, 0, na.rm = T) - 
        0.02), by = -0.02))
    df$Threshold.show <- rep(NA, nrow(df))
    for (rownum in c(2:nrow(df))) {
        if (df[rownum, 1] != df[rownum - 1, 1] | df[rownum, 2] != 
            df[rownum - 1, 2]) {
            df$Threshold.show[rownum] <- df$Threshold[rownum]
        }
    }
    bp <- ggplot(df, aes(One_minus_specificity, sensitivity, 
        label = Threshold.show))
    
        print(bp + geom_ribbon(color = "#678197", fill = "#678197", 
            aes(ymin = 0, ymax = sensitivity)) + labs(title = "ROC Curve", 
            x = "False Positive Rate", y = "True Positive Rate") + 
            annotate("text", label = paste("AUROC:", round(AREAROC, 
                4)), x = 0.55, y = 0.35, colour = "white", size = 8) + 
            theme(legend.position = "none", plot.title = element_text(size = 20, 
                colour = "#a2d2df"), axis.title.x = element_text(size = 15, 
                colour = "#a2d2df"), axis.title.y = element_text(size = 15, 
                colour = "#a2d2df")) + coord_cartesian(xlim = c(0, 
            1), ylim = c(0, 1)))
    }
   
```

### Library required packages

```{r}
library(ggplot2)
library(ggrepel)
library(caret)
library(tidyr)
library(MLmetrics) # log loss
library(dplyr)
library(viridis) # color
library(latex2exp) # math equation
```

### Import data set

```{r}
data <- read.csv("/Volumes/Viv's 4T/Glen/Phd/AI/Final/heart.csv")
```

### Data Modification

#### Change data type

```{r}
data <- data %>%
  mutate(
    sex = as.factor(sex),
    cp = as.factor(cp),
    fbs = as.factor(fbs),
    restecg = as.factor(restecg),
    exng = as.factor(exng),
    slp = as.factor(slp),
    caa = as.factor(caa),
    thall = as.factor(thall),
    output = as.factor(output)
  )
# str(data$sex) check data type
```



#### Check NA vakue

```{r}
sum(is.na(data))
```

### EDA

#### Describe data

```{r}
library(skimr)
#please try the following code by uncommenting it
skim(data)
```

#### Summary data

```{r}
summary(data)
```



#### Numeric variables

```{r}
library(plyr)
library(Rmisc)
p1 <- ggplot(data, aes(x=	trtbps, fill=output))+
  geom_histogram(binwidth = 10)+ 
  theme_light() +
  scale_fill_manual(values = c("#678197", 
                                  "#a2d2df"))+
  scale_x_continuous(name = "Resting blood pressure (in mm Hg)") +
  scale_y_continuous(name = "Count") +
    ggtitle("trtbps")+
    theme(legend.position="top", plot.title = element_text(size=10, hjust = 0.5))

p2 <- ggplot(data, aes(x=chol, fill=output))+
  geom_histogram(binwidth = 10)+ 
  theme_light() +
  scale_fill_manual(values = c("#678197", 
                                  "#a2d2df"))+
   scale_x_continuous(name = "cholestoral in mg/dl fetched via BMI sensor") +
  scale_y_continuous(name = "Count") +
  ggtitle("chol")+
  theme(legend.position="top", plot.title = element_text(size=10, hjust = 0.5))


p3 <- ggplot(data, aes(x=thalachh, fill=output))+
  geom_histogram(binwidth = 5)+ 
  theme_light() +
  scale_fill_manual(values = c("#678197", 
                                  "#a2d2df"))+
   scale_x_continuous(name = "maximum heart rate achieved") +
  scale_y_continuous(name = "Count") +
  ggtitle("thalachh")+
  theme(legend.position="top", plot.title = element_text(size=10, hjust = 0.5))


p4 <- ggplot(data, aes(x=age, fill=output))+
  geom_histogram(binwidth = 5)+ 
  theme_light() +
  scale_fill_manual(values = c("#678197", 
                                  "#a2d2df"))+
  scale_x_continuous(name = "age") +
  scale_y_continuous(name = "Count") +
  ggtitle("age")+
  theme(legend.position="top", plot.title = element_text(size=10, hjust = 0.5))


Rmisc::multiplot( p1, p2, p3,p4, cols = 2)

```

##### Remove chol outlier

```{r}
max_chol_index <- which.max(data$chol)
data<- data[-max_chol_index,]
```

```{r}
male = nrow(data[data$sex == '1' & data$output=='1',])
fe = nrow(data[data$sex == '0' & data$output=='1',])
per1 = male/nrow(data[data$sex == '1',])
per2 = fe/nrow(data[data$sex == '0',])
p3 <- ggplot(data, aes(x=sex, fill = output)) +
  geom_bar(stat = "count") +
  theme_light() +
  scale_fill_manual(values = c("#678197", 
                                  "#a2d2df"))+
                                    
   geom_text(x=1, y=fe-10, label=paste(round(per2*100),'%'), color = 'white') +
   geom_text(x=2, y=male-10, label=paste(round(per1*100),'%'), color = 'white')
                                  
          
```

#### Categorical variables

```{r}

# Function to create the bar plot with text annotations
create_bar_plot <- function(data, categorical_var, output_var) {
  # Get unique categories
  unique_categories <- unique(data[[categorical_var]])
  unique_categories <- factor(sort(as.character(unique_categories)))
  # Initialize counts and percentages vectors
  counts <- numeric(length(unique_categories))
  totals <- numeric(length(unique_categories))
  percentages <- numeric(length(unique_categories))
  
  # Calculate counts and percentages
  for (i in seq_along(unique_categories)) {
    cat <- unique_categories[i]
    counts[i] <- nrow(data[data[[categorical_var]] == cat & data[[output_var]] == '1', ])
    totals[i] <- nrow(data[data[[categorical_var]] == cat, ])
    if (counts[i] > 0) {
      percentages[i] <- counts[i] / totals[i]
    } else {
      percentages[i] <- 0
    }
  }
  

  # Create a data frame for ggplot
  plot_data <- data.frame(
    category = unique_categories,
    count = counts,
    percentage = percentages
  )

  # Create ggplot
  p <- ggplot(data, aes_string(x=categorical_var, fill=output_var)) +
    geom_bar(stat="count") +
    theme_light() +
    scale_fill_manual(values = c("#678197", "#a2d2df"))
  
  # Add text annotations
  for (i in 1:nrow(plot_data)) {
  
    if (!is.na(plot_data$percentage[i])) {
      p <- p + geom_text(
        x=i,
        y = plot_data$count[i] ,  # Adjust the y position as needed
        label = paste0(round(plot_data$percentage[i] * 100,1), '%'),
        color = 'white'
      )
    }
  }
  
  return(p)
}


```

```{r}

library(gridExtra)
# Using the function to create plots
p_sex <- create_bar_plot(data, 'sex', 'output')
p_cp <- create_bar_plot(data, 'cp', 'output')
p_fbs <- create_bar_plot(data, 'fbs', 'output')
p_res <- create_bar_plot(data, 'restecg', 'output')
p_ex <- create_bar_plot(data, 'exng', 'output')
p_slp <- create_bar_plot(data, 'slp', 'output')
p_caa <- create_bar_plot(data, 'caa', 'output')
p_tha <- create_bar_plot(data, 'thall', 'output')

# Display the plots
grid.arrange(p_sex, p_cp, p_fbs, p_res, p_ex, p_slp, p_caa,p_tha,ncol = 3)

```

### Modelling

#### Split data

(1) Training data: 85% -\> training model

(2) Validation data: 20% from training data-\> tuning model (not to specify when k-fold CV)

(3) Testing data: 15% -\> evaluation model

Note:

(1) under seed(323)

(2) Testing data us separated from the original data, and validation data is separated from training data.

```{r}
set.seed(323)

# Shuffle row indices: rows
rows <- sample(nrow(data))

# Randomly order data: fin.data
data <- data[rows, ]

# Identify row to split on (85%): split
split <- round(nrow(data) * .85)

# Create train
train_data <- data[1:split, ]

# Create val
test_data <- data[(split + 1):nrow(data), ]
```

##### Save splitted csv

```{r}
write.csv(train_data, "train_data.csv", row.names = FALSE)
write.csv(test_data, "test_data.csv", row.names = FALSE)
```

#### Logistic Regression

##### Model1: full predictor inputs
 
###### Validation data: 20% of train_data

```{r}
set.seed(323)
# Shuffle row indices: rows
rows <- sample(nrow(train_data))

# Randomly order data: fin.data
train <- train_data[rows, ]

# Identify row to split 
split <- round(nrow(train_data) * 0.8)

# Create test
val <- train[(split + 1):nrow(train), ]

# Create val
train <- train[1:split, ]

```

```{r}
# Fit logistic regression model
model_l <- glm(output ~ ., data = train, family = binomial)

# Summarize the model
result <- summary(model_l)
result
```

```{r}
# Predict on the training data
lo_val_pred_prob<- predict(model_l, newdata = val[,-c(14)],type = "response")
lo_val_pred <- as.factor(ifelse(lo_val_pred_prob > 0.5, 1, 0))

```

```{r}
actual = factor(val$output)
conf_matrix <- confusionMatrix(lo_val_pred,actual)
draw_confusion_matrix(conf_matrix)
```


##### Best subset: select variables

```{r}
library(leaps)
best = regsubsets(output~., data, nvmax = 22)
result <- summary(best)
```

###### Criterias: Cp, BIC

```{r}
Cp = result$cp
BIC = result$bic
cbind(Cp, BIC)
```

```{r}
which.min(Cp)
which.min(BIC)
```

```{r}
par(mfrow = c(1, 2))
plot(Cp, xlab = "Number of Predictors", ylab = "Cp",
     type = 'l', lwd = 2)
points(13, Cp[13], col = "pink", cex = 2, pch = 8, lwd = 2)
plot(BIC, xlab = "Number of Predictors", ylab = "BIC",
     type = 'l', lwd = 2)
points(10, BIC[10], col = "pink", cex = 2, pch = 8, lwd = 2)

```

-   Best 10 variables: sex + cp+ exng +slp +caa +thall

```{r}
coef(best, 10)
```

-   Best 13 variables: sex + cp+ + exng + slp + caa + thall + trtbps + thalachh + oldpeak

```{r}
coef(best, 13)
```

##### Best Subset with K-fold Cross Validation

Purpose: The testing data is separated from the training data, and the validation data is derived from the training data. This is done to ensure that the testing data is completely independent of the data used to train or tune the models. However, the separation of training and validation data depends on the random seed. Therefore, we use K-fold cross-validation to create different splits of the data based on various random seeds.

```{r, warning = FALSE}
 library(caret)
k = 10-5+1
# define acc matrix
acc = matrix(NA, nrow = k, ncol = 2, dimnames = list(NULL, c("lr10", "lr13")))

set.seed(323)
for(i in 1:k){
   # define training control
  train_control <- trainControl(method = "cv", number = i+4)
  
  # train the model on training set
  model_10 <- train(output ~ sex+cp+exng+slp+caa+thall,
                 data = train_data,
                 trControl = train_control,
                 method = "glm",
                 family=binomial())
  
  model_13 <- train(output ~ sex + cp+ + exng + slp + caa + thall + trtbps + thalachh + oldpeak,
                 data = train_data,
                 trControl = train_control,
                 method = "glm",
                 family=binomial())
  
  acc[i,] = c( model_10$results$Accuracy, model_13$results$Accuracy)
}

rownames(acc) <- c("k=5","k=6","k=7","k=8","k=9","k=10")
acc
```

$K=5$ and 10 variables perform best.

```{r}
colMeans(acc)
```

```{r}
ls10 <- glm(output ~ sex + cp+ + exng + slp + caa + thall, data = train_data, family = binomial)
```

So, we choose the logistics regression with 10 variables Actually, it only includes 6 variables.

##### Model2: logistic regression with 10 variables -> choose this model to test

```{r}
# Fit logistic regression model
model_2 <- glm(output ~ sex + cp+ + exng + slp + caa + thall, data = train_data, family = binomial)

# Summarize the model
result <- summary(model_2)

```

```{r}
# Predict on the training data
lo_test_pred_prob<- predict(model_2, newdata = test_data[,-c(14)],type = "response")
lo_test_pred <- as.factor(ifelse(lo_test_pred_prob > 0.5, 1, 0))
actual <- test_data$output

```

```{r}
conf_matrix <- confusionMatrix(lo_test_pred,actual)
draw_confusion_matrix(conf_matrix)
```

```{r}
ROC.plot.rv(actual,lo_test_pred_prob)
```

#### Lasso - k-fold CV

Concepts: panelty term is R1, so it can be 0 -> feature selection


```{r}

train_control <- trainControl(method = "cv", number = 5)
# Define the model formula and data
model_formula <- output ~ .

# Train the LASSO model with cross-validation
lasso_model <- train(model_formula,
                     data = train_data,
                     method = "glmnet",
                     trControl = train_control,
                     tuneGrid = expand.grid(alpha = 1,      # alpha = 1 for LASSO
                                            lambda = seq(0.001, 0.1, by = 0.001)),  # Specify a range of lambda values to search
                     family = "binomial")
plot(lasso_model)
```

```{r}
coef(lasso_model$finalModel, lasso_model$bestTune$lambda)
```



```{r}
k = 10-5+1
# define acc matrix
acc = matrix(NA, nrow = k, ncol = 2, dimnames = list(c("k=5","k=6","k=7","k=8","k=9","k=10"), c("acc","best_lambda")))

set.seed(323)
for(i in 1:k){
   # define training control
  train_control <- trainControl(method = "cv", number = i+4)
  
  model3 <- train(model_formula,
                     data = train_data,
                     method = "glmnet",
                     trControl = train_control,
                     tuneGrid = expand.grid(alpha = 1,      # alpha = 1 for LASSO
                                            lambda = seq(0.001, 0.1, by = 0.001)),  # Specify a range of lambda values to search
                     family = "binomial")
  
  
  acc[i,] = c( max(model3$results$Accuracy), model3$bestTun$lambda)
}

acc
```

We choose $\lambda=0.009$

```{r}
best_lambda <- acc[which.max(acc),2]
print(best_lambda)
```


```{r}
library(glmnet)
X <- as.matrix(train_data[, -which(names(train_data) == "output")])
y <- train_data$output
model3 <- glmnet(X, y, alpha = 1, family = "binomial", lambda = best_lambda)  
```

```{r}
# Predict on the training data
# Make predictions (response scale)
lasso_test_pred_prob <- predict(model3, newx = as.matrix(test_data[,-c(14)]), s = best_lambda, type = "response")
lasso_test_pred <- as.factor(ifelse(lo_test_pred_prob > 0.5, 1, 0))
actual <- test_data$output

```

```{r}
conf_matrix <- confusionMatrix(lasso_test_pred,actual)
draw_confusion_matrix(conf_matrix)
```

```{r}
ROC.plot.rv(actual,lasso_test_pred_prob)
```

```{r}
coef(lasso_model$finalModel, lasso_model$bestTune$lambda)
```


#### PCA -FAMD

Reference: <https://arrow.tudublin.ie/cgi/viewcontent.cgi?article=1227&context=scschcomdis>


Feature selcetion: with 13 variables(one hot encoded) as in Best Subset, but actually with 9 variables

output ~ sex + cp+ + exng + slp + caa + thall + trtbps + thalachh + oldpeak

```{r}
library("FactoMineR")
library("factoextra")
# the percentages of inertia explained by each FAMD dimensions
train.famd <- FAMD(train_data[,c(2,3,4,8,9,10,11,12,13)], graph = FALSE)

```

```{r}
eig.train <- get_eigenvalue(train.famd)
head(eig.train,10)
```

We only have 5 dimension, that $41.98%$ of explained variances.

```{r}
par(mfrow=c(2,2))
fviz_screeplot(train.famd)
# Plot of variables
fviz_famd_var(train.famd, repel = TRUE)
# Contribution to the first dimension
fviz_contrib(train.famd, "var", axes = 1)
# Contribution to the second dimension
fviz_contrib(train.famd, "var", axes = 2)
```

```{r}
library(plotly)
plot.train.famd <-as.data.frame(train.famd$ind)
x <- cbind(train_data, plot.train.famd[1:3])
## Plot
plot_ly(x, 
        x = ~coord.Dim.1, 
        y = ~coord.Dim.2, 
        z = ~coord.Dim.3, 
        color = ~output,
        marker = list(size = 7)) 
```

```{r}
# Extract train FAMD dimensions/components
train_famd_dimensions <- as.data.frame(train.famd$ind$coord)
# Combine train FAMD components with outcome variable
train_famd_data <- cbind(train_famd_dimensions, output = train_data$output)
```

```{r}
# Extract test FAMD dimensions/components
test_famd_dimensions <- as.data.frame(predict(train.famd,test_data[,-c(14)])$coord)
# Combine test FAMD components with outcome variable
test_famd_data <- cbind(test_famd_dimensions, output = test_data$output)
colnames(test_famd_data)<- c("Dim.1", "Dim.2", "Dim.3", "Dim.4", "Dim.5", "output")
```

```{r}
model_pca_glm <- glm(output ~ ., data = train_famd_data, family = "binomial")
summary(model_pca_glm)
```

```{r}
famd_test_pre_prob <- predict(model_pca_glm, test_famd_data, type = "response")
famd_test_pre <- as.factor(ifelse(famd_test_pre_prob > 0.5, 1, 0))
actual <- test_data$output
```

```{r}
conf_matrix <- confusionMatrix(famd_test_pre,actual)
draw_confusion_matrix(conf_matrix)

```

```{r}
ROC.plot.rv(actual,famd_test_pre_prob)

```

#### Tree


```{r}
library(rpart)
library(rpart.plot)
library(caret)
library(rattle)
set.seed(323)
tree <- rpart(output ~., data = train, method ='class', cp = 0.002)
fancyRpartPlot(tree)
printcp(tree)

```

```{r}
tree$variable.importance
```

```{r}
tree.pred <- predict(tree, val[,-c(14)], type = "class")
tree.pred.prob <- predict(tree, val[,-c(14)], type = "prob")[,2]
actual <- val$output
```

```{r}
conf_matrix <- confusionMatrix(tree.pred,actual)
draw_confusion_matrix(conf_matrix)
```

##### test cp

```{r}
cp <- seq(0.0001, 0.05, 0.001)
acc <-rep(0, length(cp))
set.seed(323)
for (i in 1:50){
  tree <- rpart(output ~., data = train_data, method ='class', cp = cp[i])
  tree.pred <- predict(tree, val[,-c(14)], type = "class")
  tree.pred.prob <- predict(tree, val[,-c(14)], type = "prob")[,2]
  actual <- val$output
  acc[i] <- sum(actual == tree.pred)/length(actual)
}



```

```{r}
acc_tree = data.frame(cp, acc)
ggplot(acc_tree,aes(cp, acc)) +
  geom_line() +
  theme_light()
```

Choose $cp \leq0.04$

##### important variable

```{r}
df <- data.frame(imp = tree$variable.importance)
df2 <- df %>% 
  tibble::rownames_to_column() %>% 
  dplyr::rename("variable" = rowname) %>% 
  dplyr::arrange(imp) %>%
  dplyr::mutate(variable = forcats::fct_inorder(variable))

ggplot2::ggplot(df2) +
  geom_segment(aes(x = variable, y = 0, xend = variable, yend = imp), 
               size = 1.5, alpha = 0.7) +
  geom_point(aes(x = variable, y = imp, col = variable), 
             size = 1, show.legend = F) +
  coord_flip() +
  ylab("Importance")+
  xlab("Feature")+
  theme_bw()

```

##### testing

```{r}
tree.test.pred.prob <- predict(tree, test_data[,-c(14)], type = "prob")[,2]
tree.test.pred <- as.factor(ifelse(tree.test.pred.prob > 0.5, 1, 0))

actual <- test_data$output

conf_matrix <- confusionMatrix(tree.test.pred,actual)
draw_confusion_matrix(conf_matrix)
```

##### ROC

```{r}
ROC.plot.rv(actuals =actual, predictedScores = tree.test.pred.prob)
```

#### Random forests

```{r}
library(randomForest)
```

```{r}
set.seed(323)
rf <- randomForest(output ~ ., data = train, mtry=5, ntree=1000)
rf.pred <- predict(rf, val[,-c(14)], predict.all=TRUE)
```

```{r}
err_by_tree = sapply(1:ncol(rf.pred$individual),function(i){
apply(rf.pred$individual[,1:i,drop=FALSE],1,
function(i)with(rle(i),values[which.max(lengths)]))
})

err_by_tree = colMeans(err_by_tree!=val$output)
```

```{r}
plot(rf$err.rate[,1],type="l",col="steelblue",xlab="ntrees",ylab="err.rate",
    ylim=c(0,0.5))
    lines(err_by_tree,col="orange")
    legend("topright",fill=c("steelblue","orange"),c("val","OOB.train"))
```

```{r}
set.seed(323)
rf.pred.prob <- predict(rf, val[,-c(14)], type = "prob")[,2]
rf.pred <- as.factor(ifelse(rf.pred.prob > 0.5, 1, 0))

actual <- val$output

conf_matrix <- confusionMatrix(rf.pred,actual)
draw_confusion_matrix(conf_matrix)
```

##### Tuning

```{r, warning=FALSE, message=FALSE}
library(pROC)
m <- seq(1, 50, 1)
auc <-rep(0, length(m))
set.seed(323)
for (i in 1:50){
  rf <- randomForest(output ~ ., data = train, mtry=m[i], ntree=1000)
  rf.pred <- predict(rf, val[,-c(14)], type = "class")
  rf.pred.prob <- predict(rf, val[,-c(14)], type = "prob")[,2]
  actual <- val$output
  roc<-roc(actual,rf.pred.prob)
  auc[i] <-   auc(roc)
}

```

```{r}
auc_rf = data.frame(m, auc)
ggplot(auc_rf,aes(m, auc)) +
  geom_line() +
  theme_light()
```

##### Testing

Choose $m =9$

```{r}
set.seed(323)
rf <- randomForest(output ~ ., data = train_data, mtry=9, ntree=1000)
rf.pred.prob <- predict(rf, test_data[,-c(14)], type = "prob")[,2]
rf.pred <- as.factor(ifelse(rf.pred.prob > 0.5, 1, 0))
actual <- test_data$output
```

```{r}
conf_matrix <- confusionMatrix(rf.pred,actual)
draw_confusion_matrix(conf_matrix)
ROC.plot.rv(actuals = test_data$output, predictedScores = rf.pred.prob)
```

#### XGBoost

Scaling 

```{r}
train_scaled = cbind(train_data[,-c(14)]%>%mutate_if(is.numeric,scale), output = train_data$output)
test_scaled = cbind(test_data[,-c(14)]%>%mutate_if(is.numeric,scale), output = test_data$output)
```

One-hot encode

```{r}
library(xgboost)   # for fitting the xgboost model
library(e1071)
require(Matrix)
library(caret)
sparse_matrix_train <- sparse.model.matrix(output ~ ., data = train_scaled)[,-1]
sparse_matrix_test <- sparse.model.matrix(output ~ ., data = test_scaled)[,-1]
train_output <- as.numeric(train_data$output)-1
test_output <- as.numeric(test_data$output)-1
```

```{r}
# Define parameters for xgboost
params <- list(
  objective = "binary:logistic",
  eval_metric = "auc",
  max_depth = 6,
  eta = 0.3,
  nthread = 2
)

bst <- xgboost::xgboost(data = sparse_matrix_train, label = train_output, max_depth = 4,
               eta = 1, nthread = 2, nrounds = 10,objective = "binary:logistic")

```

```{r}
importance <- xgb.importance(feature_names = colnames(sparse_matrix_train), model = bst)
head(importance)
```

- Gain is the improvement in accuracy brought by a feature to the branches it is on

- Cover measures the relative quantity of observations concerned by a feature

- Frequency is a simpler way to measure the Gain. It just counts the number of times a feature is used in all generated trees.

```{r}
xgb.plot.importance(importance_matrix = importance)
```


```{r}
# generate predictions for our held-out testing data
xg_pred_prob <- predict(bst, sparse_matrix_test)
xg.pred <- as.factor(ifelse(xg_pred_prob > 0.5, 1, 0))
conf_matrix <- confusionMatrix(xg.pred,as.factor(test_data$output))
draw_confusion_matrix(conf_matrix)
```

```{r}
# Define hyperparameters
params <- list(
  objective = "multi:softmax",
  num_class = 3,
  eval_metric = "mlogloss",
  max_depth = 6,
  eta = 0.3,
  gamma = 0,
  subsample = 1,
  colsample_bytree = 1
)
# Cross-validation to find the best number of rounds
cv <- xgb.cv(params = params, data = sparse_matrix_train,label = train_output, nrounds = 100, 
             nfold = 5, showsd = TRUE, stratified = TRUE, 
             print_every_n = 10, early_stopping_rounds = 10)

best_nrounds <- cv$best_iteration

# Train the model
model <- xgboost::xgboost(params = params, data = sparse_matrix_train,label = train_output, nrounds = best_nrounds)

```


```{r}
# generate predictions for our held-out testing data
xg_pred_prob <- predict(model, sparse_matrix_test)
xg.pred <- as.factor(ifelse(xg_pred_prob > 0.5, 1, 0))
conf_matrix <- confusionMatrix(xg.pred,as.factor(test_data$output))
draw_confusion_matrix(conf_matrix)
```

```{r}
ROC.plot.rv(actuals = test_data$output, predictedScores = xg_pred_prob)
```


Better, but still overfitting.

- Conclusion: It's well-known that a tree(/forest) algorithm (xgboost/rpart/etc.) will generally 'prefer' continuous variables over binary categorical ones in its variable selection, since it can choose the continuous split-point wherever it wants to maximize the information gain

#### Knn

```{r}
library(dplyr)
library(class)
# Initialize variables
l <- list()
maxi <- 0
ind <- 0
best <- 0
best_predict <- NULL
train_scaled = train_data[,-c(14)]%>%mutate_if(is.numeric,scale)
test_scaled = test_data[,-c(14)]%>%mutate_if(is.numeric,scale)
# Loop over a range of k values
for (i in 4:40) {
  # Train KNN model
  predict <- knn(train =train_scaled, test = test_scaled, cl = train_data$output, k = i)
  
  # Calculate accuracy
  acu <- sum(predict == test_data$output) / length(test_data$output)
  
  # Append accuracy to list
  l[[i]] <- acu
  
  # Check for the best model
  if (maxi < acu) {
    maxi <- acu
    ind <- i
    best <- acu
    best_predict <- predict
  }
}

# Print the best k and its accuracy
cat("Best k:", ind, "\n")
cat("Best auc:", best, "\n")
```



```{r}
knn_pre <- knn(train = train_scaled, test = test_scaled, cl = train_data$output, k = 14, prob = TRUE)
knn_pre_prob <-  attr(knn_pre, "prob")
conf_matrix <- confusionMatrix(knn_pre,as.factor(test_data$output))
draw_confusion_matrix(conf_matrix)
```
#### Neural Network

```{r}
library(readr)
nn.pred.prob <- read_csv("pred.csv", col_names = FALSE)
```

```{r}
ROC.plot.rv(actuals = test_data$output, predictedScores = nn.pred.prob$X1)
```


```{r}
nn.pred.prob$X1[which(nn.pred.prob$X1 > 0.5)] = "1"
nn.pred.prob$X1[which(nn.pred.prob$X1 <= 0.5)] = "0"
nn.pred.prob$X1 <- as.factor(nn.pred.prob$X1)
nn.pred <- nn.pred.prob$X1
actual <- test_data$output
library(pROC)
conf_matrix <- confusionMatrix(nn.pred,actual)
draw_confusion_matrix(conf_matrix)
```